Distortion corrected kernel density estimate on Riemannian manifolds

### 2022-10-30 paper/paper.Rmd
Sec 4&5 done

<!-- Anomaly detection has been an important and diverse area where anomalies or outliers are detected in a given data. It often involves a data analysis process to uncover the unusual patterns and has been widely applied to machine learning [@Omar2013-lz], network intrusions identification [@Ahmed2016-co; @Bhuyan2013-ra], medical imaging [@Fernando2022-ef], fraudulent attacks [@Ahmed2016-dz], cyber-security [@Ten2011-co], energy consumption [@Cheng2021-ex], the last of which explored detecting households with anomalous electricity usage distributions instead of raw data. In the case of non-Euclidean sample space, the observations lie on a Riemannian manifold embedded in a very high-dimensional ambient space, which makes it computationally expensive or impossible to detect anomalies. To address this problem, we propose a kernel density estimator of the low-dimensional statistical manifold embedding and find outliers as the data with the lowest density estimates. -->

@@

For samples points lying on a manifold with the differentiable structure called the Riemannian manifold, @Pelletier2005-vu generalize the kernel density estimator based on the kernel weights from the geodesic distance between the estimated points and the sample points. The idea of the estimator is to use a strictly positive function of the geodesic distance on the manifold and then normalize it with the volume density function of the Riemannian manifold for curvature [@Henry2009-ll]. 
However, in many application scenarios, we tend to find that the sample points are not drawn directly from the manifolds because they are embedded in a much higher-dimensional space. Therefore, the kernel density estimator from @Pelletier2005-vu is not applicable because the geodesic distance and the volume density function are unknown.
This is when we introduce manifold learning to reduce the input data dimension. For these high-dimensional data set, various manifold learning algorithms including ISOMAP, LLE, Laplacian Eigenmaps, t-SNE, and UMAP (see details of these algorithms in @Cheng2021-ex), could be applied to get a low-dimensional embedding, which are used as approximations of the underlying manifold.

@@

By exploiting the connection between the estimated Riemannian metric and the Riemannian geodesic distance as well as the volume density function for curvature, we propose the main contribution of the paper, which is the variable distortion-corrected kernel density estimator (DC-KDE) for manifold learning embedding. Starting from the high-dimensional sample data, we apply manifold learning algorithms to get the low-dimensional embedding in the same dimensional space as the underlying manifold together with the estimated Riemannian matrix at each embedding point. Then the DC-KDE is used to estimate the density of the manifold and distortions induced by manifold learning methods are fixed with the estimated geometric information. Our distortion-corrected estimator is novel in filling the gap between the high-dimensional sample space and the density of the unknown manifold. These density estimates are useful in many other areas, including classification, clustering and anomaly detection. Similar to @Cheng2021-ex, the highest density region plots[@Hyndman1996-lk] could be generated using the kernel density estimates for outlier visualization, which brings a novel anomaly detection method for Riemannian manifolds.

@@

One limitation of the kernel density estimator is that the density can be estimated only at points where data have been observed since the estimator requires the Riemannian $\pmb{H}_j$. To estimate the density at points that do not belong to the sampled points, any smoothed average of nearest neighbors can be used instead. We note that the particular downstream task that we are interested in is anomaly detection for which only the density estimates at observed sample points are required because anomalies are identified as the points with lowest density. The entire workflow is summarized in \autoref{fig:vkde}. The last two steps in \autoref{fig:vkde} are our main contributions, generating distortion-corrected KDE with adaptive Riemannian metric $\pmb{H}_i$ at each point and computing the highest density region plots based on the density estimates for anomaly detection. Compared to the anomaly detection with a general kernel density estimator in @Cheng2021-ex, the changes are also highlighted in blue. With this anomaly detection process, outliers based on lowest densities could be detected more accurately regardless of the distortion in manifold learning.

@@

## Irish smart meter dataset

In this application, we use the smart meter data from the *CER Smart Metering Project - Electricity Customer Behaviour Trial, 2009-2010* in Ireland [@cer2012-data] between 14 July 2009 and 31 December 2010. The CER dataset[^2] records the half-hourly electricity consumption of individual residential and commercial properties, not including energy for cooling or heating systems. We selected the $3,639$ residential data with no missing values during the data collection period for a total of
$535$ days.

@@

For the electricity consumption data of residential individuals, it would be worthwhile to explore the distribution of electricity demand rather than the raw consumption data to study the usage patterns of different households or different periods or the week [@Hyndman2018-ia].
@Cheng2021-ex propose two non-Euclidean distance estimators to enable manifold learning algorithms in statistical manifolds with each observation as a distribution. @Cheng2021-ex use the same smart meter data for identifying outliers with kernel bandwidth estimation but fail to consider the distortion and information loss in the 2-dimensional embeddings given that the input data dimension is much higher. By introducing the Riemannian metric as the bandwidth matrix, we could take into account the distortion in the 2-D embedding and further improve the accuracy of the density estimation.

In this section, we first calculated the same empirical distributions of the $336$ half-hourly periods of the week for each household, and apply the total variation distance estimator proposed in @Cheng2021-ex in the statistical manifold learning to get the 2-D embedding of all households. Equation \@ref(eq:denestimator) is then used to obtain density estimates with the pointwise variable Riemannian metric as the bandwidth matrix and detect outliers. The data processing steps have been clearly
stated in the application section of @Cheng2021-ex and they are skipped here. Unlike the simulations in \autoref{simulation}, we know nothing about the true density of the electricity distributions for all periods of the week and all households, so it is impossible to compare the estimated densities with the true meta data density as in \autoref{fig:fivedisomapden}. However, we could generate all the density estimates with the existing KDE method with fixed bandwidth, which is an optimal method for density estimation, and compare the densities from our proposed method with them.

```{r electricityoutliers, fig.align = 'center', out.width = "90%", fig.cap="Electricity usage plots of all 535 days for the most typical household and two anomalies in rows and two bandwidth selection methods in columns.", echo = FALSE, eval=T}
knitr::include_graphics("figures/kde_2densities_raw_electricity.png")
```

\autoref{fig:electricityoutliers} shows the electricity usage data of three households for both density estimation methods respectively, with the top one being the most typical household with the highest density and the bottom two being the top two outliers with the lowest densities.
The typical households in the top row are close except that there are a few spikes for the one with fixed bandwidth. As to the anomalies, variable bandwidth tends to capture the unusual electricity demand volume when the usage is very low or high. It could also capture the unusual usage pattern when there are sudden spikes in ID 3243 or very high base electricity usage for the middle time periods in ID 2838. In contrast, fixed bandwidth KDE is more sensitive to spikes even when the spikes happen in a certain time window in 7049, or when the usage has an obvious time-of-week pattern with a few low electricity usage periods.

```{r electricityhdrsfixed, fig.align = 'center', out.width = "90%", fig.cap="Two smart-meter demand examples, ID 1003 and ID 1539, from the Irish smart meter data set.", echo = FALSE, eval=T}
knitr::include_graphics("figures/hdr_electricity_hdrbox_3id_7dow.png")
```

```{r electricityhdrsvkde, fig.align = 'center', out.width = "90%", fig.cap="Two smart-meter demand examples, ID 1003 and ID 1539, from the Irish smart meter data set.", echo = FALSE, eval=T}
knitr::include_graphics("figures/electricity_hdrbox_3id_7dow.png")
```

Further insights could be gained by comparing the quantile region plots of electricity demand against the time of the week for the same typical or anomalous households in \autoref{fig:electricityhdrsfixed} and \autoref{fig:electricityhdrsvkde}. Again the distribution of both typical households in the top panel has shown a repeated period-of-the-week usage pattern, with higher usage during mealtime on all seven days of the week and slightly higher usage for weekends.
However, this repeated pattern in a week window is clearly for the typical household ID 1321. As for the distributions for outliers, the middle row outliers from variable bandwidth have spikes only on Tuesday and Sunday noons, while the fixed bandwidth has an increasing electricity demand across the day of the week. The bottom row outliers both have a repeated time usage pattern, but the electricity usage amount is higher with the highest median above 3kWh. These findings show the difference in finding typical and anomalous households with different bandwidth selections.

@@

In this paper, we propose a new method to estimate the density of manifold learning embedding and further identify outliers based on the densities. The Riemannian metric measure the direction and angle of the distortion when mapping data points through a non-linear function in manifold learning algorithms. By introducing the Riemannian metric as the pointwise variable bandwidth matrix in kernel density estimation, the local distortion in the low-dimensional embedding could be used to estimate densities, leading to a more accurate description of the data distributions. We compare our proposed method with fixed bandwidth KDE by two simulation settings, 2-D meta data mapped as a 3-D swiss roll or twin peaks data and 5-D semi-hypersphere mapped in 100-D space, and show that variable bandwidth could improve the density estimation given a good manifold learning embedding.

As an empirical example, we explore the distributions of different households and time periods of the week in the Irish smart meter data. Five manifold learning algorithms, including ISOMAP, LLE, Laplacian Eigenmaps, t-SNE, and UMAP, are applied to get the 2-D embeddings, and KDE with both variable and fixed bandwidth are used to get the density estimates. We compare both density estimates by looking at the distributions of the most typical households with the highest densities and the most anomalous households with the lowest densities. Both methods could identify the typical households with certain usage patterns, while the outliers are anomalous in different ways.

There are several open questions to be explored. The first involves the selection of tuning parameters for the manifold algorithm so that a maximal level of embedding quality is achieved, where embedding quality is measured using one of the metrics discussed in the online supplementary material of @Cheng2021-ex. The scaling of the Riemannian metric to get the closest range of the true densities is also worth exploring. The scale of distortion in each point in manifold learning could vary a lot. If we could smooth across all data points, eg. multiply the Riemannian matrix with the ratio between its determinant
and the sum of all Riemannian matrix determinants, the global density estimates could be potentially smoothed. The density estimates on the edges of the whole data structure could be improved because most outer area points tend to be detected as outliers. The choice of manifold learning algorithms also has a large impact on the embedding accuracy, which we have seen will affect the density estimation and outlier detection. However, the outperformance of VKDE with Riemannian bandwidth than the fixed bandwidth has been shown in the higher dimensional simulation data and the electricity usage data, which are more related
to real-life data sets.

<!-- Shortcoming and other references for future direction (improvement on density estimation, other kernels, densities on edges, bad embeddings with inaccurate distortions, multi-dimensional kernel density estimate). -->

<!-- Discuss simulation with non-uniform meta data. -->

<!-- The aim of the project is not to evaluate different manifold learning algorithms.  Although ISOMAP worked OK here it is still distorted and in real examples, you will never 'know' that the data are uniform on the sphere.  Instead, the idea of this project is to do kernel density estimation in a way that takes distortion into account.
-->

@@

# Appendix: Simulation with swiss roll mapping {#swissrollappe}

@@

To further compare the accuracy of the estimated densities for all data points, we calculate the correlation between the rank of true densities and the estimated densities and present in \autoref{tab:srcors}. It can be seen that the rank correlation of our proposed method with variable bandwidth is higher for LLE and UMAP, although the correlation for UMAP is very close to zero. The highest correlation comes from our method in LLE embedding, which is mainly due to it being closest to the
rectangular structure of the meta data shown in \autoref{fig:metadensity}. For Laplacian Eigenmaps, our method has wrongly estimated the left area with lower densities even though their true densities are very high in yellow, leading to a negative
correlation. The negative correlation would occur typically when the highest or lowest true density areas are not well estimated. As for the estimates in highly distorted embedding, including ISOMAP, t-SNE, and UMAP, the rank correlations are quite low. This shows that our proposed method could improve the kernel density estimate of manifold learning embedding by considering the distortion using the Riemannian metric.
However, if the distortion is too severe, eg. ISOMAP, or when the embedding is discontinuous, eg. t-SNE and UMAP, the density estimates are not as reliable for outlier detection.

<!-- Since there are four kernels in the meta data and darker points are the defined outliers, we could derive that for ISOMAP and LLE, variable kernel density estimates could detect the number of kernels more accurately than the KDE with fixed bandwidth. As for t-SNE and UMAP, both embeddings are quite close to each other with points clustered together, but the mixture of kernel structure is not as clear as LLE. Instead, we could roughly compare the location of outliers. For t-SNE, the middle plot has indexed more true outliers than the bottom plot; while for UMAP, both the middle and bottom plot fail to find the true outliers, but the outliers from fixed bandwidth are more scattered along with the curved embedding. -->

>>>>

<!-- Anomaly detection has been an important and diverse area where anomalies or outliers are detected in a given data. It often involves a data analysis process to uncover the unusual patterns and has been widely applied to machine learning [@Omar2013-lz], network intrusions identification [@Ahmed2016-co; @Bhuyan2013-ra], medical imaging [@Fernando2022-ef], fraudulent attacks [@Ahmed2016-dz], cyber-security [@Ten2011-co], energy consumption [@Cheng2021-dh], the last of which explored detecting households with anomalous electricity usage distributions instead of raw data. In the case of non-Euclidean sample space, the observations lie on a Riemannian manifold embedded in a very high-dimensional ambient space, which makes it computationally expensive or impossible to detect anomalies. To address this problem, we propose a kernel density estimator of the low-dimensional statistical manifold embedding and find outliers as the data with the lowest density estimates. -->

@@

For samples points lying on a manifold with the differentiable structure called the Riemannian manifold, @Pelletier2005-vu generalize the kernel density estimator based on the kernel weights from the geodesic distance between the estimated points and the sample points. The idea of the estimator is to use a strictly positive function of the geodesic distance on the manifold and then normalize it with the volume density function of the Riemannian manifold for curvature [@Henry2009-ll]. 
However, in many application scenarios, we tend to find that the sample points are not drawn directly from the manifolds because they are embedded in a much higher-dimensional space. Therefore, the kernel density estimator from @Pelletier2005-vu is not applicable because the geodesic distance and the volume density function are unknown.
This is when we introduce manifold learning to reduce the input data dimension. For these high-dimensional data set, various manifold learning algorithms including ISOMAP, LLE, Laplacian Eigenmaps, t-SNE, and UMAP (see details of these algorithms in @Cheng2021-dh), could be applied to get a low-dimensional embedding, which are used as approximations of the underlying manifold.

@@

By exploiting the connection between the estimated Riemannian metric and the Riemannian geodesic distance as well as the volume density function for curvature, we propose the main contribution of the paper, which is the variable distortion-corrected kernel density estimator (DC-KDE) for manifold learning embedding. Starting from the high-dimensional sample data, we apply manifold learning algorithms to get the low-dimensional embedding in the same dimensional space as the underlying manifold together with the estimated Riemannian matrix at each embedding point. Then the DC-KDE is used to estimate the density of the manifold and distortions induced by manifold learning methods are fixed with the estimated geometric information. Our distortion-corrected estimator is novel in filling the gap between the high-dimensional sample space and the density of the unknown manifold. These density estimates are useful in many other areas, including classification, clustering and anomaly detection. Similar to @Cheng2021-dh, the highest density region plots[@Hyndman1996-lk] could be generated using the kernel density estimates for outlier visualization, which brings a novel anomaly detection method for Riemannian manifolds.

@@

One limitation of the kernel density estimator is that the density can be estimated only at points where data have been observed since the estimator requires the Riemannian $\pmb{H}_j$. To estimate the density at points that do not belong to the sampled points, any smoothed average of nearest neighbors can be used instead. We note that the particular downstream task that we are interested in is anomaly detection for which only the density estimates at observed sample points are required because anomalies are identified as the points with lowest density. The entire workflow is summarized in \autoref{fig:vkde}. The last two steps in \autoref{fig:vkde} are our main contributions, generating distortion-corrected KDE with adaptive Riemannian metric $\pmb{H}_i$ at each point and computing the highest density region plots based on the density estimates for anomaly detection. Compared to the anomaly detection with a general kernel density estimator in @Cheng2021-dh, the changes are also highlighted in blue. With this anomaly detection process, outliers based on lowest densities could be detected more accurately regardless of the distortion in manifold learning.

@@

<!-- ## Irish smart meter dataset -->

In this application, we use the smart meter data from the *CER Smart Metering Project - Electricity Customer Behaviour Trial, 2009-2010* in Ireland [@cer2012-data] between 14 July 2009 and 31 December 2010. The CER dataset[^2] records the half-hourly electricity consumption of individual residential and commercial properties, but not including energy for cooling or heating systems. We selected the $3,639$ residential data with no missing values during the data collection period for a total of $535$ days.

@@

For the electricity consumption data of residential individuals, it would be worthwhile to explore the distribution of electricity demand rather than the raw consumption data to study the usage patterns of different households or different periods or the week [@Hyndman2018-ia].
@Cheng2021-dh propose two non-Euclidean distance estimators to be used in manifold learning algorithms in case of statistical manifolds with each observation as a distribution. The same Irish smart meter data is used for identifying outliers with kernel density estimation, but they fail to consider the distortion and information loss in the 2-dimensional embeddings. Therefore, we propose in this paper to apply distortion corrected KDE for anomaly detection as the downstream task of dimension reduction. Again, we use ISOMAP, LLE, Laplacian Eigenmaps, t-SNE and UMAP for dimension reduction to get the 2-D embedding for kernel density estimation and anomaly detection. To this end, we compare the density estimates from KDE with and without distortion correction and show how robust the anomalies are to different dimension reduction algorithms with the use of distortion correction. We use the highest density regions plots to visualize the density estimates. However, for this real data set with unknown structure, the ground truth densities are unknown and it is not possible to tell which anomalies are the true ones. Therefore, we will further examine the distributions of the anomalous households to study the typical or anomalous electricity usage patterns.

Although full details for data processing and dimension reduction are provided in @Cheng2021-dh, we briefly describe the procedures here. First, for each household and each of the $336$ half-hourly periods of the week, we build the empirical distributions and apply the total variation distance estimator for statistical manifold learning, which gives a 2-dimensional embedding for each dimension reduction method. Then we apply the kernel density estimator with distortion corrected for each method to the low-dimensional embedding and compare it with the KDE results in the Section 4.4 of @Cheng2021-dh. The bandwidth parameter $r$ is set as 1 in Equation \@ref(eq:denestimator).

 ```{r electricityembedding, fig.align = 'center', out.width = "90%", fig.cap="Highest density region plots of the 2-D embedding for $3,639$ households, with each point representing the distribution of one household and colors indicating the density estimated from DC-KDE (left panel) and KDE (right panel). The black points with blue text indexing are the top 20 anomalous households found by each dimension reduction method in each row.", echo=FALSE, eval=F}
# knitr::include_graphics("figures/")
```

<!-- The results are summarized in \autoref{fig:electricityembedding}.
TODO -->

<!-- In this section, we first calculated the same empirical distributions of the $336$ half-hourly periods of the week for each household, and apply the total variation distance estimator proposed in @Cheng2021-dh in the statistical manifold learning to get the 2-D embedding of all households. Equation \@ref(eq:denestimator) is then used to obtain density estimates with the pointwise variable Riemannian metric as the bandwidth matrix and detect outliers. The data processing steps have been clearly
stated in the application section of @Cheng2021-dh and they are skipped here. Unlike the simulations in \autoref{simulation}, we know nothing about the true density of the electricity distributions for all periods of the week and all households, so it is impossible to compare the estimated densities with the true meta data density as in \autoref{fig:fivedisomapden}. However, we could generate all the density estimates with the existing KDE method with fixed bandwidth, which is an optimal method for density estimation, and compare the densities from our proposed method with them.
 -->
```{r electricityoutliers, fig.align = 'center', out.width = "90%", fig.cap="Electricity usage plots of all 535 days for the most typical household and two anomalies in rows and KDE with distortion correction (left panel) and KDE (right panel) in columns.", echo = FALSE, eval=T}
knitr::include_graphics("figures/kde_2densities_raw_electricity.png")
```

\autoref{fig:electricityoutliers} shows the electricity usage data of three households for both density estimation methods respectively, with the top one being the most typical household with the highest density and the bottom two being the top two outliers with the lowest densities.
The typical households in the top row are similar except that there are a few spikes for household $1472$ when using KDE. As for the anomalies, distortion corrected KDE tends to capture the unusual electricity demand when the usage is very low or high in volume. It could also capture the unusual usage pattern when there are sudden spikes in ID 3243 or very high base electricity usage for the middle time periods in ID 2838. In contrast, KDE without distortion correction is more sensitive to spikes even when the spikes happen in a certain time window in 7049, or when the usage has an obvious time-of-week pattern with a few low electricity usage periods.

```{r electricityhdrsfixed, fig.align = 'center', out.width = "90%", fig.cap="Quantile region plots of electricity demand against the time of week for one typical household 1321 and two anomalies, 3243 and 2838. The quantile regions displayed in the plot are 99.0\\%, 95.0\\%, 75\\%, and 50\\%. Household 1321 has a repeated period-of-the-week electricity usage pattern slightly different for weekdays and weekends.", echo = FALSE, eval=T}
knitr::include_graphics("figures/hdr_electricity_hdrbox_3id_7dow.png")
```

```{r electricityhdrsvkde, fig.align = 'center', out.width = "90%", fig.cap="Quantile region plots of electricity demand against the time of week for one typical household 1472 and two anomalies, 7049 and 5136. The day-of-the-week patterns is much more different for Household 7049 compared with 5136, the latter of which does not have the obvious meal time related usage patterens.", echo = FALSE, eval=T}
knitr::include_graphics("figures/electricity_hdrbox_3id_7dow.png")
```

Further insights could be gained by comparing the quantile region plots of electricity demand against the time of the week for the same typical or anomalous households in \autoref{fig:electricityhdrsfixed} and \autoref{fig:electricityhdrsvkde}. Again the distribution of both typical households in the top panel has shown a repeated period-of-the-week usage pattern, with higher usage during mealtime on all seven days of the week and slightly higher usage for weekends.
This repeated pattern in a week window is also obvious for the typical household ID 1321 from DC-KDE. As for the distributions for outliers, the middle row outliers from variable bandwidth have spikes only on Tuesday and Sunday noons, while the fixed bandwidth has an increasing electricity demand across the day of the week. The bottom row outliers both have a repeated time usage pattern, but the electricity usage amount is higher with the highest median above 3kWh. These findings show the difference in finding typical and anomalous households with different kernel density estimation methods.

@@

In this paper, we propose a novel distortion correction method to estimate the density of an embedding from manifold learning algorithms and further identify outliers based on the densities. Compared with KDE, our distortion corrected KDE makes use of geometric information for each data point to correct the distortion induced by the embedding. The Riemannian metric is estimated with the Learn Metric algorithm to approximate the geodesic distance and volume density function locally at each point. We compare our proposed method with KDE by two simulation settings, a 2-D manifold embedded as a 3-D twin peaks shape and a 4-D manifold mapped in a 100-D ambient space, and show that DC-KDE could generate more accurate kernel density estimation is more robust to the choice of dimension redution algorithm.

As an empirical example, we explore the distributions of $3,639$ households and $336$ time periods of the week in the Irish smart meter data. Five manifold learning algorithms, including ISOMAP, LLE, Laplacian Eigenmaps, t-SNE, and UMAP, are applied to get the 2-D embedding, followed by density estimation with KDE and DC-KDE. Without the ground truth density, we compare both density estimates by looking at the distributions of the most typical households with the highest densities and the anomalous households with the top two lowest densities. Both methods could identify the typical households with certain usage patterns, while the outliers are anomalous in different ways.

There are several open questions to be explored. The first involves the selection of bandwidth parameters for kernel density estimation, which has been explored in many KDE-related literatures. The second one is related to the quality of the embedding from dimension reduction methods. Although our distortion correction method is fairly robust to different choices of manifold learning methods, in certain cases, when the data structure is too complex and the distortion is too severe to correct, the quantitive relationship between embedding quality and density estimation accuracy is not immediate clear. The embedding quality could be measured using one of the metrics discussed in the online supplementary material of @Cheng2021-dh, but when the ground truth densities are unknown, which is usually the case with real-world data set, it is hard to tell wherther the distortion are corrected in the right way. The density estimates on the edges of the whole data structure could alos be explored because most outer area points tend to be detected as outliers. However, the outperformance of DC-KDE than KDE has been shown in the higher dimensional simulation data and the electricity usage data, which are more related to real-life data sets.

<!-- kernel types discussion? -->

@@

<!-- # Appendix: Simulation with swiss roll mapping {#swissrollappe}

@@

To further compare the accuracy of the estimated densities for all data points, we calculate the correlation between the rank of true densities and the estimated densities and present in \autoref{tab:srcors}. It can be seen that the rank correlation of our proposed method with variable bandwidth is higher for LLE and UMAP, although the correlation for UMAP is very close to zero. The highest correlation comes from our method in LLE embedding, which is mainly due to it being closest to the
rectangular structure of the meta data shown in \autoref{fig:metadensity}. For Laplacian Eigenmaps, our method has wrongly estimated the left area with lower densities even though their true densities are very high in yellow, leading to a negative
correlation. The negative correlation would occur typically when the highest or lowest true density areas are not well estimated. As for the estimates in highly distorted embedding, including ISOMAP, t-SNE, and UMAP, the rank correlations are quite low. This shows that our proposed method could improve the kernel density estimate of manifold learning embedding by considering the distortion using the Riemannian metric.
However, if the distortion is too severe, eg. ISOMAP, or when the embedding is discontinuous, eg. t-SNE and UMAP, the density estimates are not as reliable for outlier detection.
 -->
<!-- Since there are four kernels in the meta data and darker points are the defined outliers, we could derive that for ISOMAP and LLE, variable kernel density estimates could detect the number of kernels more accurately than the KDE with fixed bandwidth. As for t-SNE and UMAP, both embeddings are quite close to each other with points clustered together, but the mixture of kernel structure is not as clear as LLE. Instead, we could roughly compare the location of outliers. For t-SNE, the middle plot has indexed more true outliers than the bottom plot; while for UMAP, both the middle and bottom plot fail to find the true outliers, but the outliers from fixed bandwidth are more scattered along with the curved embedding. -->
